\section{Introduction}
\label{sec:introduction}

Scientific simulation, computer graphics, and collision detection as well as other fields utilize structures called octrees. An octree allows a three dimensional space to be broken down hierarchically into a tree structure with up to eight children per node. Meshless point clouds, such as a group of particles, are an important application of such structures, especially when paired with Barnes-Hut or Fast Multipole Methods which allow reductions in the runtime of force calculations between points from $O(n^2)$ to $O(nlog(n))$ and roughly $O(n)$ respectively. This has driven the development of octree construction methods that operate on GPUs and across multiple nodes in clusters and supercomputers. 

One of the most interesting new methods for meshless point clouds is Cornerstone \cite{cornerstone}. In the case of simulating the movement of particles in dynamic systems, the octree must be recalculated at each time step, however the primary advantage of Cornerstone is to rebalance the tree rather than rebuild from scratch. Cornerstone has been heavily used on exascale supercomputers, however not all problems which may benefit from this method require such systems. Furthermore a profile of the code internals and how the method performs across varying conditions of particle distributions, parameters, and other variants does not appear publicly but may help to optimize and evidence the methods usefulness across varying systems and problem spaces. This project seeks to do such a characterization and will focus closely on the mapping of Cornerstone to the current GPU architectures examining the kernels as well as the communication cost of doing these calculations across nodes.
