\section{Remaining Work}

There is significant potential to expand the breadth of our project over the second half of the semester. First of all as we mentioned there is at least one opportunity which is clear for optimization using kernel fusion and/or stream processing of the radix sort and gather of particle properties, dependent on profiling results. A kernel fusion would reduce global memory pressure by not writing the sorted data to global memory, then launching another kernel and reading it back into the register file. We also have consistent access to a wider variety of hardware now including A100s, A40s, and H200s. We will want to gather new metrics for those types of GPUs and in the configurations available and scale out our LET testing to more GPUs. It is possible this will provide unexpected results that require a hardware and kernel comparison. In general we will also be diving to kernel level for some of the bottleneck kernels, for example if we develop a fused kernel and look at cache performance, kernel metrics like occupancy and resource usage. We will also gather more data on the overwhelming bottleneck for calculating the LET which was presented in \ref{sec:results} of the calculation of peer ranks, then we will do a more through breakdown of the communication cost of both the domain decomposition and LET calculation.

An interesting parameter to study would be $L$ and $\theta$, which is the maximal depth an octree can be resolved to and the accuracy parameter for the MAC respectively. The two values of $L$ usually used are 32-bit integers which allow resolution to a depth of 10 levels and 64-bit integers which allow resolution to a depth of 21. The tradeoff can sacrifice the capability to resolve to greater depths that are necessary in very large cases but will reduce the amount of global memory pressure. This is especially of interest since the radix sort and gather kernels are bottlenecks in the local case both of which will be memory bound kernels that are performance capped by memory throughput. The $\theta$ parameter may be of interest because as we saw in preliminary data the thing we believe is capping performance is peer discovery and $\theta$ reduces the number of peers.

So far we have tested mostly uniform perturbations, we plan to augment this to test many different types of affine transforms that the method may encounter such as rotations and translations. We also want to look at domain over-sizing, the domain bounding box defines the total domain of the particles. If this box is significantly larger than the true domain of the particles it will cause the octree to have large regions of empty space which requires the octree to become deeper thus inducing inefficiencies. A study on the relative volume of the bounding box compared to the volume of say a unit cube against the build time should be conducted. 

In summary our action plan for the coming weeks is:
\begin{enumerate}
    \item Optimization and profiling cycle to optimize the local case.
    \item Look at the low level kernel data to study any cache effects, kernel parameters, or resource details.
    \item Study new node configurations and gather more data on the domain decomposition and LETs, produce in depth metrics on the MPI communication cost.
    \item Do a parameter sweep for different values of $\theta$ and $L$, generate metrics similar to those presented here. If the reduction in global memory pressure caused by using $L = 10$ (a 32-bit integer) rather than $L = 21$ (a 64-bit integer) provides significant gains do a study on the roofline chart and pressure on the memory subsystem, along with comments on the reduction in utility of the octree due to the reduction in maximal octree depth.
    \item Test affine transforms for perturbations and the effect of domain over-sizing.
\end{enumerate}
